### 2013-04-04 Sunpy_paper_0.3.tex
Added template for IOP CSD.
This is the jouranl to which scipy 2013 are submitting, however
if we want to / need to at a later date this can be changed.

\documentclass[12pt]{iopart}

%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\begin{document}

\title{SunPy - Python for Solar Physicists}

\author{SunPy Developers}

\address{The Internet}
\ead{sunpy@googlegroups.com}
\begin{abstract}

SunPy contains too much awesome.

\end{abstract}

\maketitle

\section{Introduction}

\section*{References}
\begin{thebibliography}{10}
\bibitem{book1} Goosens M, Rahtz S and Mittelbach F 1997 {\it The \LaTeX\ Graphics Companion\/} 
(Reading, MA: Addison-Wesley)
\bibitem{eps} Reckdahl K 1997 {\it Using Imported Graphics in \LaTeX\ } (search CTAN for the file `epslatex.pdf')
\end{thebibliography}

\end{document}