### 2014-01-22 Sunpy_paper_0.4.tex

\documentclass[12pt]{iopart}

\usepackage{minted}
\usepackage{hyperref}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\usepackage{graphicx}
\graphicspath{{./images/}}

\begin{document}

\title{SunPy - Python for Solar Physics}

\author{The SunPy Community}
\address{\url{http://sunpy.org}}
\ead{sunpy@googlegroups.com}

\author{S J Mumford}
\address{Solar Physics \& Space Plasma Research Centre (SP$^{2}$RC), School of 
Mathematics and Statistics, The University of Sheffield, Hicks Building, 
Hounsfield Road, Sheffield, S3 7RH U.K.}

\begin{abstract}
This paper presents version 0.4 of SunPy a community developed Python package 
for solar physics.

\end{abstract}

\maketitle

\input{1-Introduction}
\label{sec:Intro}

\input{2-Data_Types}
\label{sec:DataTypes}

\input{3-Retrieval}
\label{sec:retrevial}

\input{4-Util}
\label{sec:util}

\input{5-Dev}
\label{sec:dev}

\input{6-Future}
\label{sec:future}

%%% References
\bibliography{sunpy_paper_0.4}{}

\end{document}