### 2014-01-28 Sunpy_paper_0.4.tex
Some small typo fixed and latex stuff

\documentclass[12pt]{iopart}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\usepackage{natbib}
\newcommand*{\newblock}{Bibliography}

\usepackage{minted}
\definecolor{bg}{rgb}{0.95, 0.95, 0.95}
\usepackage{hyperref}

\usepackage{graphicx}
\graphicspath{{./images/}}

\usepackage{perpage} %the perpage package
\MakePerPage{footnote} %the perpage package command

\begin{document}

\title{SunPy - Python for Solar Physics}

\input{authors}

\input{abstract}

\maketitle

\input{1-Introduction}

\input{2-Data_Types}

\input{3-Retrieval}

\input{4-Util}

\input{solarsoft}

\input{5-Dev}

\input{6-Future}

%%% References
\bibliographystyle{stunat}
\bibliography{Sunpy_paper_0.4}{}

\end{document}