### 2014-01-28 3-Retrieval.tex
minor typo corrections. Added a brief introduction to Section 3 (data search and retrieval)

\section{Solar Data Search and Retrieval}\label{sec:retrieval}

Several well-developed resources currently exist within heliophysics which provide event databases, remote access and data retrieval for a large number of data sources. As a community-based project, SunPy provides support for these resources via the \texttt{sunpy.net} package. In the following subsections, we describe a number of these resources and introduce current methods for accessing them within SunPy.

\input{3-1-VSO}
\input{3-2-HEK}
\input{3-3-HELIO}
\input{3-4-Helioviewer}
\input{3-5-Database}