### 2014-01-30 3-Retrieval.tex

\section{Solar Data Search and Retrieval}\label{sec:retrieval}

Several well-developed resources currently exist within heliophysics which provide  
remote access and data retrieval for a large number of data sources as well as access to 
event databases. As a community-based project, 
SunPy provides support for these resources via the \texttt{net} subpackage. 
In the following subsections, we describe a number of these resources and 
how to use them in SunPy.

\input{3-1-VSO}
\input{3-2-HEK}
\input{3-3-HELIO}
\input{3-4-Helioviewer}
\input{3-5-Database}