### 2014-01-30 authors.tex
Added Michael affiliation and British spelling to the abs. Notice Summary.tex is missing

\author{The SunPy Community,}
\address{\url{http://sunpy.org}}
\ead{sunpy@googlegroups.com}

\author{
% People who have directly contributed to the text, still in no order.
Stuart J Mumford$^1$,
Steven Christe$^2$,
David P\'erez-Su\'arez$^3$,
Andrew R Inglis$^{2,5}$,
Jack Ireland$^{6}$,
Albert Y Shih$^2$,
Simon Liedtke$^{11}$,
Russell J Hewett$^8$, %88
% Contributors to SunPy Sorted thus:
% No. Commits to main repo + commits in open PR's
% Date of most recent contribution
Florian Mayer$^{12}$, %721
Keith Hughitt$^4$, %446
Nabil Freij$^1$, %52
Matthew Earnshaw, %52
Tomas Meszaros$^{10}$, %17
N Chorley, %17
Samuel M Bennett$^1$, %14
M Malocha$^{16}$, %10
John Evans, %10 %No affil
Ankit Agrawal$^{15}$, %8
M Bates, %6
Andrew J Leonard$^7$, %3
J Ballew, %3
Thomas P Robitaille$^9$, %2
Benjamin Mampaey$^{13}$, %2
Jose Iv\'an Campos-Rozo$^{14}$, %1
Michael S. Kirk$^{2}$
and 
A Kustov %1
}

\address{
$^1$ Solar Physics \& Space Plasma Research Centre (SP$^{2}$RC), 
School of Mathematics and Statistics, The University of Sheffield, Hicks 
Building, Hounsfield Road, Sheffield, S3 7RH U.K.

$^2$NASA Goddard Space Flight Center, Greenbelt, MD, USA

$^3$South African National Space Agency - Space Science, Hospital 
Street, 7200 Hermanus, Western Cape, South Africa

$^4$Department of Cell Biology and Molecular Genetics, University of Maryland, College Park

$^5$ The Catholic University of America, Washington, DC, USA

$^6$ADNET Systems Inc., Mail Code 671.1, NASA Goddard Space Flight Center, 
Greenbelt, MD, USA

$^7$Department of Mathematics and Physics, Aberystwyth University, 
Physical Sciences Building, Aberystwyth, Ceredigion, SY23 3BZ

$^8$Department of Mathematics, Massachusetts Institute of Technology, 77 
Massachusetts Ave, E17-317, Cambridge, MA 02142

$^9$ Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg
69117, Germany

$^{10}$ Masaryk University, Faculty of Informatics, Botanicka 68a Brno

$^{11}$ University of Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany

$^{12}$ Vienna University of Technology, Karlsplatz 13, 1040 Vienna, Austria

$^{13}$ Royal Observatory of Belgium

$^{14}$ Observatorio Astronómico Nacional Universidad Nacional de Colombia 
Bogotá, D.C., Colombia

$^{15}$ Indian Institute of Technology, Bombay

$^{16}$ Humboldt State University, 1 Harpst St, Arcata, CA 95521, United States
}