### 2014-01-30 3-Retrieval.tex

\section{Solar Data Search and Retrieval}\label{sec:retrieval}

Several well-developed resources currently exist which provide remote access to 
and data retrieval form a large number of solar and heliospheric data sources 
and event databases. SunPy provides support for these resources via the 
\texttt{net} subpackage. In the following subsections, we describe a number of 
these resources and how to use them in SunPy.

\input{3-1-VSO}
\input{3-2-HEK}
\input{3-3-HELIO}
\input{3-4-Helioviewer}
\input{3-5-Database}