### 2014-01-31 authors.tex

\author{The SunPy Community,}
\address{\url{http://sunpy.org}}
\ead{sunpy@googlegroups.com}

\author{
% People who have directly contributed to the text, still in no order.
Stuart J Mumford$^1$,
Steven Christe$^2$,
David P\'erez-Su\'arez$^3$,
Jack Ireland$^{2,4}$,
Albert Y Shih$^2$,
Andrew R Inglis$^{2,5}$,
Simon Liedtke$^{6}$,
Russell J Hewett$^7$, %88
% Contributors to SunPy Sorted thus:
% No. Commits to main repo + commits in open PR's
% Date of most recent contribution
Florian Mayer$^{8}$, %721
Keith Hughitt$^9$, %446
Nabil Freij$^1$, %52
Tomas Meszaros$^{10}$, %17
Samuel M Bennett$^1$, %14
Michael Malocha$^{11}$, %10
John Evans$^{12}$,%10 %No affil
Ankit Agrawal$^{13}$, %8
Andrew J Leonard$^{14}$, %3
Thomas P Robitaille$^{15}$, %2
Benjamin Mampaey$^{16}$, %2
Jose Iv\'an Campos-Rozo$^{17}$, %1
and Michael S Kirk$^{2}$
}

\address{
$^1$Solar Physics \& Space Plasma Research Centre (SP$^{2}$RC), 
School of Mathematics and Statistics, The University of Sheffield, Hicks 
Building, Hounsfield Road, Sheffield, S3 7RH, UK

$^2$NASA Goddard Space Flight Center, Greenbelt, MD, USA

$^3$South African National Space Agency - Space Science, Hospital 
Street, 7200 Hermanus, Western Cape, South Africa

$^4$ADNET Systems Inc., Mail Code 671.1, NASA Goddard Space Flight Center, 
Greenbelt, MD, USA

$^5$The Catholic University of America, Washington, DC, USA

$^{6}$University of Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany

$^7$Department of Mathematics, Massachusetts Institute of Technology, 77 
Massachusetts Ave, E17-317, Cambridge, MA, USA

$^{8}$Vienna University of Technology, Karlsplatz 13, 1040 Vienna, Austria

$^9$Department of Cell Biology and Molecular Genetics, University of Maryland, College Park, MD, USA

$^{10}$Masaryk University, Faculty of Informatics, Botanicka 68a Brno, Czech Republic

$^{11}$Humboldt State University, 1 Harpst St, Arcata, CA, USA

$^{12}$Boston Python User Group, Boston, MA, USA

$^{13}$Indian Institute of Technology, Bombay, India

$^{14}$Department of Mathematics and Physics, Aberystwyth University, 
Physical Sciences Building, Aberystwyth, SY23 3BZ, UK

$^{15}$Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg
69117, Germany

$^{16}$Royal Observatory of Belgium, Brussels, Belgium

$^{17}$Observatorio Astron\'omico Nacional Universidad Nacional de Colombia 
Bogotá, D.C., Colombia

}