### 2014-01-31 Summary.tex

\section{Summary}
We have presented the release of SunPy (v0.4), a Python package for solar physics. In
this paper we have described the main functionality which includes the SunPy data types, 
\texttt{Map} (see Section~\ref{ssec:map}), \texttt{Lightcurve} (see Section~\ref{ssec:lightcurve}), and \texttt{Spectrogram} (see Section~\ref{sec:spectra}).
We have described the data and event catalogue retrieval capabilities of SunPy for
the Virtual Solar Observatory (see Section~\ref{ssec:vso}), the Heliophysics Event Knowledgebase (see Section~\ref{ssec:hek}), as well as
the Heliophysics Integrated Observatory (see Section~\ref{ssec:helio}). We described
a new organization tool for data files integrated into SunPy (see Section~\ref{ssec:db}) 
and we discussed the community aspects, development model (see Section~\ref{sec:dev}), and future plans (see Section~\ref{sec:future}) for the project.
We invite members of community to contribute to the effort by using SunPy for their 
research, reporting bugs, and sharing new functionality with the project.