### 2014-01-31 2-Data_Types.tex

\section{Core Data Types}\label{sec:DataTypes}

The core of SunPy are a set of data strucutures that are specifically designed for the
three primary varieties of solar physics data: images, time series, and
spectra. These core data types are supported by SunPy's classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series),
and \texttt{Spectrum} and \texttt{Spectrogram} (1D and 2D spectra).

These classes allow access to the data
and associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute, while the metadata should be stored 
in the \texttt{meta} attribute (currently, only \texttt{Map} has this feature 
fully implemented). 
% schriste - change this until all fixed
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, and arrays.  
In order to provide instrument-specific specialisation, the core SunPy classes 
make use of subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the \textit{SDO}/AIA instrument.

All of the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

This design philosophy makes the behaviour of SunPy's visualisation 
routines intuitive to those who already understand the \texttt{matplotlib}
interface, as well as allowing the use of the standard 
\texttt{matplotlib} commands to manipulate the plot parameters (e.g., title, axes).
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control.

This section will give a brief overview of the \textit{current} functionality 
of each of the core SunPy data types.

\input{2-1-Map}
\input{2-2-Lightcurve}
\input{2-3-Spectra}