### 2014-01-31 acknowledgements.tex

\section{Acknowledgements}
Many of the larger features in SunPy have been developed with the generous support of 
external organizations. 
Initial development of SunPy's VSO and HEK implementations were funded by ESA's 
Summer of Code In Space (SOCIS) program, as well as a prototype GUI and an
N-dimensional data-type implementation. In 2013, with support from Google's 
Summer Of Code (GSOC) program, through the Python Software Foundation, the 
\texttt{helio}, \texttt{hek2vso}, and \texttt{database} subpackages were 
developed. The Spectra and Spectrogram classes were implemented with support 
from the Astrophysics Research Group at Trinity College Dublin, Ireland, in 
2012.