### 2014-08-25 Sunpy_paper_0.4.tex

\documentclass[12pt]{iopart}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\usepackage{natbib}
\newcommand*{\newblock}{Bibliography}

\usepackage[section]{minted}
%\usepackage{upquote}
%\AtBeginDocument{%
%\def\PYZsq{\textquotesingle}%
%}
\definecolor{bg}{rgb}{0.95, 0.95, 0.95}

\newmintedfile[pythoncode]{pycon}{
bgcolor=bg,
fontfamily=tt,
linenos=true,
numberblanklines=true,
numbersep=12pt,
numbersep=5pt,
gobble=0,
frame=leftline,
framerule=0.4pt,
framesep=2mm,
funcnamehighlighting=true,
tabsize=4,
obeytabs=false,
mathescape=false
samepage=false, %with this setting you can force the list to appear on the same page
showspaces=false,
showtabs =false,
texcl=false,
}

\usepackage{hyperref}

\usepackage{graphicx}
\graphicspath{{./images/}}

\usepackage{perpage} %the perpage package
\MakePerPage{footnote} %the perpage package command

\usepackage{aas_macros}

\begin{document}

\title{SunPy - Python for Solar Physics}

\input{authors}

\input{abstract}

\maketitle

\input{1-Introduction}

\input{2-Data_Types}

\input{3-Retrieval}

\input{4-Util}

\input{5-Dev}

\input{6-Future}

\input{Summary}

\input{acknowledgements}

%%% References
\bibliographystyle{stunat}
\bibliography{Sunpy_paper_0.4,refs}{}

\end{document}