### 2015-05-11 Sunpy_paper_0.4.tex
Add option to include code using listings.

\documentclass[12pt]{iopart}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\usepackage{natbib}
\newcommand*{\newblock}{Bibliography}

% You can use either minted or listings to include the code 
% in this file. listings will hopefully work on arxiv.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% Use Minted % % % % % % % % % % % % % % % % % % % % % % % 
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%\usepackage[section]{minted}

%\newmintedfile[pythoncode]{pycon}{
%fontfamily=tt,
%linenos=true,
%numberblanklines=true,
%numbersep=12pt,
%numbersep=5pt,
%gobble=0,
%frame=leftline,
%framerule=0.4pt,
%framesep=2mm,
%funcnamehighlighting=true,
%tabsize=4,
%obeytabs=false,
%mathescape=false
%samepage=false, %with this setting you can force the list to appear on the same page
%showspaces=false,
%showtabs =false,
%texcl=false,
%}

% % % % % % % % % % % % % % % % % % % % %
% Use Listings % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % 
\usepackage{listings}
\usepackage{float}
\newfloat{listing}{thHp}{lop}
\renewcommand{\pythoncode}{\lstinputlisting}

% % % % % % % % % % % % % % % % % % % %

\usepackage{hyperref}

\usepackage{graphicx}
\graphicspath{{./images/}}

\usepackage{perpage} %the perpage package
\MakePerPage{footnote} %the perpage package command

\usepackage{aas_macros}

\begin{document}

\title{SunPy - Python for Solar Physics}

\input{authors}

\input{abstract}

\maketitle

\input{1-Introduction}

\input{2-Data_Types}

\input{3-Retrieval}

\input{4-Util}

\input{5-Dev}

\input{6-Future}

\input{Summary}

\input{acknowledgements}

%%% References
\bibliographystyle{stunat}
\bibliography{Sunpy_paper_0.4}{}

\end{document}